Parallelizing AdaBoost and Decision Trees

### 2015-08-25 paper.tex

Since the tree-building process is the most time-consuming portion of AdaBoost,
our goal is to parallelize this. It seems natural to divide amongst processors
the work of finding the dimension in the feature space that best splits the
region and increases the purity for each node in the tree. The datasets we are
handling have several dimensions, on the order of tens to hundreds, so each
processor could be responsible for a group of dimensions in the feature space.
After each processor determines the best dimension to split amongst its
group, we can perform a reduction and determine the best split over all
dimensions.

>>>>

Since the tree-building process is the most time-consuming portion of AdaBoost,
our goal is to parallelize this. It seems natural to divide amongst processors
the work of finding the dimension in the feature space that best splits the
region and increases the purity for each node in the tree. The datasets we are
handling have several dimensions, on the order of tens to hundreds, so each
processor could be responsible for a group of dimensions in the feature space.

After each processor determines the best dimension to split on amongst those
it's responsible for, we can perform a reduction and determine the best split
over all dimensions. Any tree will split the feature space multiple times so
there are several occasions for parallelism to speed up execution. In testing I
commonly set the maximum depth for a tree to between two to five levels. Also
the number of rounds of AdaBoost, and hence the number of trees constructed, is
a parameter in the method allowing you to affect how often the algorithm drops
into parallel structure.

I've noticed with the MNIST I can achieve a speedup by a factor of around 4
when using four processors, yet increasing beyond this does not give any
noticeable gains.

Executing this parallelized tree building
